Palmer Archipelago (Antarctica) penguins: an alternative to Edgar Anderson's iris data

### 2020-06-19 manuscript/manuscript_text.Rmd

```{r setup, include=FALSE}
library(tidyverse)
library(palmerpenguins)
library(kableExtra)
library(patchwork)
library(ggExtra)
library(gt)
library(ggpubr)
library(ggbeeswarm)
knitr::opts_chunk$set(echo = TRUE)
```

@@

Allison Horst, Alison Hill, Kristen Gorman

@@

In 1936, one year after its publication in the *Bulletin of the American Iris Society*, Anderson’s *Iris* measurements were further analyzed and published by eugenicist and statistician Ronald A. Fisher in *The Annals of Eugenics* [@fisher]. Regardless of Anderson's initial motivation, the dataset today inextricably linked to Fisher’s eugenics research and even commonly (if unfairly) referred to as "*Fisher's* iris data" (ref). Due to its history in - and association with - Fisher's eugenics research, there is momentum to replace Iris with alternative datasets that offer similar teaching and learning opportunities.

@@

- Available by an appropriate license, ideally CC0 (no rights reserved)
- Should feature intuitive subjects and variables that are understandable to learners across fields
- Dimensions should be manageable (but not trivial)
- Contain real-world observations for improved understanding and engagement [@cafe]
- Provide similar opportunities for teaching data science skills and statistical methods

@@

As in `iris`, multiple opportunities for comparisons between groups also exist in `penguins`.

@@

iris_bee <- ggplot(iris, aes(x = Species, y = Petal.Width)) +
  geom_boxplot(color = "gray80",
               fill = "gray92", 
               outlier.color = NA
               ) +
  geom_beeswarm(aes(color = Species), 
                alpha = 0.6,
                size = 2) + 
  scale_color_manual(values = c("darkorange","darkorchid","cyan4")) +
  theme_minimal() +
  theme(legend.position = c(0.2, 0.8),
        plot.title.position = "plot") +
  labs(x = "Iris species",
       y = "Petal width (cm)",
       title = "Iris petal width")

fpenguins_bee <- penguins %>% 
  ggplot(aes(x = species, y = flipper_length_mm)) +
  geom_boxplot(color = "gray80",
               fill = "gray92", 
               outlier.color = NA
               ) +
  geom_beeswarm(aes(color = species), 
                alpha = 0.6,
                size = 2) + 
  scale_color_manual(values = c("darkorange","darkorchid","cyan4")) +
  theme_minimal() +
  theme(legend.position = c(0.2, 0.8),
        plot.title.position = "plot") +
  labs(x = "Penguin species",
       y = "Flipper length (mm)",
       title = "Penguin flipper length")

fpenguins_bee + iris_bee
```

@@

#### Comparison 3: Principal components analysis

>>>>

```{r setup, include=FALSE}
library(tidyverse)
library(here)
library(palmerpenguins)
library(kableExtra)
library(patchwork)
library(ggExtra)
library(gt)
library(ggpubr)
library(ggbeeswarm)
knitr::opts_chunk$set(echo = TRUE)
```

@@

Allison Horst, Alison Hill, Kristen Gorman (order TBD)

@@

In 1936, one year after publication in the *Bulletin of the American Iris Society*, Anderson’s *Iris* measurements were further analyzed and published by eugenicist and statistician Ronald A. Fisher in *The Annals of Eugenics* [@fisher]. Regardless of Anderson's initial motivation, the dataset today inextricably linked to Fisher’s eugenics research and even commonly (if unfairly) referred to as "*Fisher's* iris data" (ref). Due to its history in - and association with - Fisher's eugenics research, there is momentum to replace Iris with alternative datasets that offer similar teaching and learning opportunities.

@@

- Available by an appropriate license, ideally CC0 (no rights reserved)
- Feature intuitive subjects and variables understandable to learners across fields
- Manageable, but not trivial, size
- Real-world observations for improved understanding and engagement [@cafe]
- Provide similar opportunities for teaching data science skills and statistical methods

@@

As in `iris`, multiple opportunities for comparisons between groups also exist in `penguins` (Figure X). Notably, more obvious separation between species exists in `iris` (Figure X E & F), compared to any between-species comparisons in `penguins` (Figure X A - D).

@@

# Update this with an actual theme FFS and use '&' in patchwork

iris_pw <- ggplot(iris, aes(x = Species, y = Petal.Width)) +
  geom_boxplot(aes(color = Species),
               show.legend = FALSE) +
  scale_color_manual(values = c("darkorange","darkorchid","cyan4")) +
  theme_bw() +
  theme(plot.title.position = "plot") +
  labs(x = "",
       y = "Petal width (cm)") +
  coord_flip()

iris_pl <- ggplot(iris, aes(x = Species, y = Petal.Length)) +
  geom_boxplot(aes(color = Species),
               show.legend = FALSE) +
  scale_color_manual(values = c("darkorange","darkorchid","cyan4")) +
  theme_bw() +
  theme(plot.title.position = "plot") +
  labs(x = "",
       y = "Petal length (cm)") +
  coord_flip()

iris_sl <- ggplot(iris, aes(x = Species, y = Sepal.Length)) +
  geom_boxplot(aes(color = Species),
               show.legend = FALSE) +
  scale_color_manual(values = c("darkorange","darkorchid","cyan4")) +
  theme_bw() +
  theme(plot.title.position = "plot") +
  labs(x = "",
       y = "Sepal length (cm)") +
  coord_flip()

iris_sw <- ggplot(iris, aes(x = Species, y = Sepal.Width)) +
  geom_boxplot(aes(color = Species),
               show.legend = FALSE) +
  scale_color_manual(values = c("darkorange","darkorchid","cyan4")) +
  theme_bw() +
  theme(plot.title.position = "plot") +
  labs(x = "",
       y = "Sepal width (cm)") +
  coord_flip()

penguins_flip <- penguins %>%
  ggplot(aes(x = species, y = flipper_length_mm)) +
  geom_boxplot(aes(color = species),
               show.legend = FALSE) +
  scale_color_manual(values = c("darkorange","darkorchid","cyan4")) +
  theme_bw() +
  theme(plot.title.position = "plot") +
  labs(x = "",
       y = "Flipper length (mm)") +
  coord_flip()

penguins_mass <- penguins %>%
  ggplot(aes(x = species, y = body_mass_g)) +
  geom_boxplot(aes(color = species),
               show.legend = FALSE) +
  scale_color_manual(values = c("darkorange","darkorchid","cyan4")) +
  theme_bw() +
  theme(plot.title.position = "plot") +
  labs(x = "",
       y = "Body mass (g)") +
  coord_flip()

penguins_bill_length <- penguins %>%
  ggplot(aes(x = species, y = bill_length_mm)) +
  geom_boxplot(aes(color = species),
               show.legend = FALSE) +
  scale_color_manual(values = c("darkorange","darkorchid","cyan4")) +
  theme_bw() +
  theme(plot.title.position = "plot") +
  labs(x = "",
       y = "Bill length (mm)") +
  coord_flip()

penguins_bill_depth <- penguins %>%
  ggplot(aes(x = species, y = bill_depth_mm)) +
  geom_boxplot(aes(color = species),
               show.legend = FALSE) +
  scale_color_manual(values = c("darkorange","darkorchid","cyan4")) +
  theme_bw() +
  theme(plot.title.position = "plot") +
  labs(x = "",
       y = "Bill depth (mm)") +
  coord_flip()

penguins_box <- penguins_flip / penguins_mass / penguins_bill_length / penguins_bill_depth + plot_layout(guides = "collect")

iris_box <- iris_pw / iris_pl / iris_sw / iris_sl + plot_layout(guides = "collect")

penguins_box | iris_box + plot_annotation(
  title = 'Penguin and iris sizes',
  subtitle = 'Penguin size measurements (left) and iris petal and sepal dimensions (right)'
)

ggsave(here("figures","penguins_iris_box.png"), height = 7, width = 8)

```

### Classification?

### Other comparative analyses?

## Selected activities

### Hypothesis testing

- t-tests
- ANOVAs
- MANOVAs

### Linear regression

- Application, diagnostics, interpretation & predictions

## Random other stuff

@@

## Differences by islands?

```{r}
# No difference in Adelie across islands: Chinstrap only on Dream, Gentoo only on Torgerson
penguins %>% 
  ggplot(aes(x = island, y = body_mass_g)) +
  geom_point() +
  facet_wrap(~species)

penguins %>% 
  ggplot(aes(x = island, y = flipper_length_mm)) +
  geom_point() +
  facet_wrap(~species)

penguins %>% 
  ggplot(aes(x = island, y = bill_length_mm)) +
  geom_point() +
  facet_wrap(~species)

penguins %>% 
  ggplot(aes(x = island, y = bill_depth_mm)) +
  geom_point() +
  facet_wrap(~species)
```
